Proof No Diamond (3 layers)

### 2023-10-20 Paper/Paper.tex
test

\begin{definition}[Sign of Node]
	We define a sign function $S$ on nodes $n$ such that: 	(1). $S(n)=0$ if all path from $n$ to $c$ has 0 weight; (2). $S(n)=1$ if all path from $n$ to $c$ has non-negative weight, and at least one path has a positive weight; (3). $S(n)=-1$ if all path from $n$ to $c$ has non-positive weight, and at least one path has a negative weight. 
	In general, $S$ may not be defined on every node (e.g. if there is a negative and a positive path from $a$ to $c$). However, if there is no Diamond, $S$ is defined on all nodes: any node fulfills one of above cases (1),(2),(3).

@@

\end{proof}

@@

The DeepPoly function for $b$ can be expressed by the following steps.
First, we partition the set of indices $i$ as $I,I_0,I^+,I^-$ with:
\begin{itemize}
	\item $i \in I$ if the sign of $b_i^-=\sum_{j} w_{a_j b_i} a_j + B_{b_i}$ is always positive irrespective of $Y=(a_j)_j$, in which case 
	$\bar{f}(b_i^+)(Y)=\underline{f}(b_i^+)=\sum_{j} w_{a_j b_i} a_j + B_{b_i}$, and else:
	\item $i \in I_0$ if the sign of $b_i^-=\sum_{j} w_{a_j b_i} a_j + B_{b_i}$ is always negative irrespective of $Y=(a_j)_j$, in which case $\bar{f}(b_i^+)=\underline{f}(b_i^+)=0$,	and else:
	\item $i \in I^+$ if $w_{b_i c}>0$, and:
	\item $i \in I^-$ if $w_{b_i c}<0$.
	\end{itemize}

Remark that for all $i \notin I$, we have $\underline{f_i}(b_i^+)(Y)=0$.

We have:
\begin{align*}
	\bar{f}(c)(Y) = &
	bias_c + \sum_{b \in I} w_{bc}\sum_{a} w_{a b} weight(a) + bias_{b} \\
	& + \sum_{b \in I^+} w_{b c} \cdot \bar{f}(ReLU(\sum_{a} w_{a b} weight(a) 
	+ Bias_{b}))(Y) 
	\end{align*}

>>>>

\begin{definition}[Sign of Node]\label{sign_of_nodes}
	We define a sign function $S$ on nodes $n$ such that: 	(1). $S(n)=0$ if all path from $n$ to $c$ has 0 weight; (2). $S(n)=1$ if all path from $n$ to $c$ has non-negative weight, and at least one path has a positive weight; (3). $S(n)=-1$ if all path from $n$ to $c$ has non-positive weight, and at least one path has a negative weight. 
	In general, $S$ may not be defined on every node (e.g. if there is a negative and a positive path from $a$ to $c$). However, if there is no Diamond, $S$ is defined on all nodes: any node fulfills one of above cases (1),(2),(3).

@@

	Similarly, 
	if $\bar{f}(n^-)=max(n^-)<0$, 
	then $\bar{f}(n^+)=max(n^+)=0$, 
	and otherwise 
	$\bar{f}(n^+)=max(n^+)=\bar{f}(n^-)=bar(n^-)$.

	Hence we have equality before and after ReLU in all cases.
\end{proof}

\section{No Diamond theorem part 2}

Now we try to show the part 2 of No Diamond theorem. Again, we begin with 3 layers with only one output node $z$.

\begin{theorem}[No Diamond theorem, part 2] \label{no_diamond_2}
	Using MILP method, if all nodes $b_j$ such that there is a compensating path pair $\pi,\pi'$ with $\pi = a_k,b_j,z$ are opened, then the upper and lower bounds computed by MILP are the exact max and min value of $z$.
\end{theorem}

With out loss of generality, we only need to show the max side. The min side is almost the same.

\begin{definition}
	Let $K$ be the set of all input nodes $a_k$. We define a decomposition $K=I\sqcup J$ as follows:

	An input node $a_k$ is in $I$ if for any two paths $\pi,\pi'$ that begin with $a_k$ (end with $z$), they are not compensating, i.e., the weights of them are both non negative or both non positive; hence all path $\pi$ that begin with $a_k$ (end with $z$) are non negative (or non positive).

	An input node $a_k$ is in $J$ if it is not in $I$. We use $a_i$ to denote input nodes in $I$ and use $a_j$ to denote input nodes in $J$.
\end{definition}

\begin{lemma}
	A node $b_i$ in the hidden layer will not be opened in Theorem \ref{no_diamond_2} iff one of the following two happen:

	1. $w_{zb_i}=0$.

	2. For any input nodes $a_j\in J$, $w_{b_ia_j}=0$.
\end{lemma}

\begin{proof}
	First, we show if one of 1,2 happens, then $b_i$ will not be opened. If $w_{zb_i}=0$, it is obvious.

	If 2 happens, suppose $\pi,\pi'$ is a pair of compensating path, then $\pi = a_k,b_i,z$. Because this is a compensating path, $w_{b_ia_k}\neq 0$. But because of 2, this means that $a_k$ is in $I$, and hence there is a compensating path begin with $a_k\in I$, a contradiction. Therefore, $b_i$ cannot be a node in a compensating pair.

	Second, we show if both 1,2 not happen, then $b_i$ will be opened. Because of negation 2, there is a node $a_j$ in $J$ such that $w_{a_jb_i}\neq 0$, and since $w_{zb_i}\neq 0$, $\pi=a_j,b_i,z$ is a path with a  nonzero weight. By symmetry, we assume, $w_{zb_i}w_{b_ia_j}>0$.

	Now, since $a_j$ is in $J$, there are two paths $\pi',\pi''$ that begin with $a_j$ but $\pi''$ has a positive weight and $\pi'$ has a negative weight. Consider $\pi'$. We claim that $\pi,\pi'$ is a compensating pair that make $b_i$ opened. $\pi'$ cannot be $a_j,b_i,z=\pi$ because $\pi$ has a positive weight. So $\pi,\pi'$ are two different paths with opposite signs of values.

	So, $b_i$ will be opened. Now we prove this lemma. 
\end{proof}

Because for input nodes in $I$, they also do not have compensating path pair, we can define the sign as in Definition \ref{sign_of_nodes}:

\begin{definition}[Sign of Nodes in $I$]\label{sign_of_nodes_in_I}
	We define a sign function $S$ on $I$ such that: 	(1). $S(n)=0$ if all path from $n$ to $z$ has 0 weight; (2). $S(n)=1$ if all path from $n$ to $z$ has non-negative weight, and at least one path has a positive weight; (3). $S(n)=-1$ if all path from $n$ to $z$ has non-positive weight, and at least one path has a negative weight.

	For an input node $a_i$ in $I$:
	\begin{enumerate}
		\item  $S(a_i)=0$ if 
		for all $b_j, w(a_i b_j)\cdot w(b_j c) = 0$

		\item $S(a_i)=-1$ if for all $b_j\ w(a_i b_j)\cdot w(b_j c) \leq 0$ and there exists 
		$j, w(a_i b_j)\cdot w(b_j c) < 0$ 
	\end{enumerate}
\end{definition}

\begin{lemma}
	node $z$ can get its maximal value if for all $a_i$ in $I$, $a_i$ gets its maximal if $S(a_i)=1$ or $a_i$ gets its minimal if $S(a_i)=-1$.
\end{lemma}

\begin{proof}

	We use $a_S$ where $S\subseteq K$ to refer the input vector $\langle a_k\rangle_{a_k\in S}$. Then $a_I\oplus a_J = a_K$ and $c=c(a_K)=c(a_I,a_J)$.

	This is because \begin{align*}
		\max_{a_K} c(a_K) = \max_{a_J}\max_{a_I} c(a_I,a_J)
	\end{align*}

	For any fixed input of $a_J$, $c(\cdot, a_J)$ can be regarded as a DNN $D'$ with input nodes $a_i\in I$ by regarding all $a_J$ and their propagation in hidden layers as bias. In the simplified $D'$, there is no compensating path because of the definition of $I$. Therefore we can apply the No Diamond theorem, part 1 to get that: (for fixed $a_J$) $c(\cdot,a_J)$ can get its maximal value if for all $a_i$ in $I$, $a_i$ gets its maximal if $S(a_i)=1$ or $a_i$ gets its minimal if $S(a_i)=-1$.

	If we use $\max(\tilde{a}_I$ to denote above values, then we have $$\max_{a_I}c(a_I,a_J)=c(\max(\tilde{a}_I),a_J).$$ And therefore, $\max c=\max_{a_J} c(\max\tilde{a}_I,a_J)$
\end{proof}

Because for any method, like MILP, the upper bound computed will not be smaller that the actual maximal value, so we only need to show, $\mathrm{UB}_{\mathrm{MILP}}\leq \max c$ in any case.

Suppose $O$ is the set of nodes opened. Then for any $b_h\notin O$, $b_h=\sum_{a_i\in I}w_{b_ha_i}+B_h$ ($B_h$ is the bias). Then for any subdomain $d$ in one of  $2^{|O|}$ subdomains, suppose we compute an upper bound when $a_J=\bar{a}_J$ and $a_I=\bar{a}_I$, we show that this upper bound does not exceed the maximal value of $z$.

By the algorithm, $$\mathrm{UB} = B_z+\sum_{b_h\in O}w_{ab_h}\ReLU(b_h(\bar{a}_I,\bar{a}_J))+$$

@@

%
%The DeepPoly function for $b$ can be expressed by the following steps.
%First, we partition the set of indices $i$ as $I,I_0,I^+,I^-$ with:
%\begin{itemize}
%	\item $i \in I$ if the sign of $b_i^-=\sum_{j} w_{a_j b_i} a_j + B_{b_i}$ is always positive irrespective of $Y=(a_j)_j$, in which case 
%	$\bar{f}(b_i^+)(Y)=\underline{f}(b_i^+)=\sum_{j} w_{a_j b_i} a_j + B_{b_i}$, and else:
%	\item $i \in I_0$ if the sign of $b_i^-=\sum_{j} w_{a_j b_i} a_j + B_{b_i}$ is always negative irrespective of $Y=(a_j)_j$, in which case $\bar{f}(b_i^+)=\underline{f}(b_i^+)=0$,	and else:
%	\item $i \in I^+$ if $w_{b_i c}>0$, and:
%	\item $i \in I^-$ if $w_{b_i c}<0$.
%	\end{itemize}
%	
%Remark that for all $i \notin I$, we have $\underline{f_i}(b_i^+)(Y)=0$.
%	
%
%We have:
%\begin{align*}
%	\bar{f}(c)(Y) = &
%	bias_c + \sum_{b \in I} w_{bc}\sum_{a} w_{a b} weight(a) + bias_{b} \\
%	& + \sum_{b \in I^+} w_{b c} \cdot \bar{f}(ReLU(\sum_{a} w_{a b} weight(a) 
%	+ Bias_{b}))(Y) 
%	\end{align*}
